---
title: Analysis of Financial Statements
author: Colin James
date: '2020-07-03'
slug: analysis-of-financial-statements
categories: []
tags: []
subtitle: ''
summary: ''
authors: []
lastmod: '2020-07-03T15:00:32-04:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r Installing Necessary Packages, message=FALSE, warning=FALSE}
install.packages("XBRL")
install.packages("finreportr")

library("ggplot2")
library("quantmod")
library("dplyr")
library("finreportr")
library("XBRL")

```
## Analysis of Financial Statments

In this blog we are going to connect to the EDGAR database and analyze the most recent 10Ks of Apple

##Liquidity Ratios

#XBRL

Quick note about the finreportr function.  XBRl stands for eXtensive Business Reporting Language is a global framework to exchange business information.  The reason this R package makes this information so readily accessible is because the package is able to convert XBRL into a more usable format.  

```{r Current Ratio}
#install the financial report (finreportr) R package. 
  CompanyInfo("AAPL")
  
  AnnualReports("AAPL")
  
  appleincomestatement <- GetIncome("AAPL", 2010)
  
#Liquidity Ratios, Current and Quick.  Current Ratio = Current Assets/Current Liabilities.
  
#Current Ratio
applebalancesheet10 <- GetBalanceSheet("AAPL", 2010)
Year_2010 <- as.numeric(applebalancesheet10[31,3])/as.numeric(applebalancesheet10[49,3])

applebalancesheet11 <- GetBalanceSheet("AAPL", 2011)
Year_2011 <- as.numeric(applebalancesheet11[31,3])/as.numeric(applebalancesheet11[49,3])

applebalancesheet12 <- GetBalanceSheet("AAPL", 2012)
Year_2012 <- as.numeric(applebalancesheet12[28,3])/as.numeric(applebalancesheet12[48,3])

applebalancesheet13 <- GetBalanceSheet("AAPL", 2013)
Year_2013 <- as.numeric(applebalancesheet13[30,3])/as.numeric(applebalancesheet13[50,3])

applebalancesheet14 <- GetBalanceSheet("AAPL", 2014)
Year_2014 <- as.numeric(applebalancesheet14[30,3])/as.numeric(applebalancesheet14[52,3])

applebalancesheet15 <- GetBalanceSheet("AAPL", 2015)
Year_2015 <- as.numeric(applebalancesheet15[30,3])/as.numeric(applebalancesheet15[54,3])

applebalancesheet16 <- GetBalanceSheet("AAPL", 2016)
Year_2016 <- as.numeric(applebalancesheet16[16,3])/as.numeric(applebalancesheet16[42,3])

applebalancesheet17 <- GetBalanceSheet("AAPL", 2017)
Year_2017 <- as.numeric(applebalancesheet17[16,3])/as.numeric(applebalancesheet17[40,3])

#Create table for 
C_Ratios <- as.data.frame(rbind(Year_2010,Year_2011,Year_2012,Year_2013, Year_2014,Year_2015,Year_2016, Year_2017))

C_Years <- rbind("2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017")

CA_By_Year <- as.data.frame(cbind(C_Years, C_Ratios))

#Created a simple line graph to show the Current Ratio Over time
ggplot(data = CA_By_Year, aes(x = C_Years, y = V1, group=1)) + 
  geom_line(color="red") +
  geom_hline(aes(yintercept = mean(V1)), color="blue") +
  labs(title="Current Ratios for Apple", x="Year", y="Current Ratio")

```


```{r Quick Ratio}
#install the edgar R package
  appleincomestatement <- GetIncome("AAPL", 2010)
  
#Liquidity Ratios, Current and Quick.
#Quick Ratio = Current Assets - Inventories/Current Liabilities
  
#Quick Ratio
Year_2010 <- (as.numeric(applebalancesheet10[31,3])-as.numeric(applebalancesheet10[23,3]))/as.numeric(applebalancesheet10[49,3])

Year_2011 <- (as.numeric(applebalancesheet11[31,3])-as.numeric(applebalancesheet11[23,3]))/as.numeric(applebalancesheet11[49,3])

Year_2012 <- (as.numeric(applebalancesheet12[30,3])-as.numeric(applebalancesheet12[22,3]))/as.numeric(applebalancesheet12[48,3])

Year_2013 <- (as.numeric(applebalancesheet13[30,3])-as.numeric(applebalancesheet13[22,3]))/as.numeric(applebalancesheet13[50,3])

Year_2014 <- (as.numeric(applebalancesheet14[30,3])-as.numeric(applebalancesheet14[22,3]))/as.numeric(applebalancesheet14[52,3])

Year_2015 <- (as.numeric(applebalancesheet15[30,3])-as.numeric(applebalancesheet15[22,3]))/as.numeric(applebalancesheet15[54,3])

Year_2016 <- (as.numeric(applebalancesheet16[16,3])-as.numeric(applebalancesheet16[10,3]))/as.numeric(applebalancesheet16[42,3])

Year_2017 <- (as.numeric(applebalancesheet17[16,3])-as.numeric(applebalancesheet17[10,3]))/as.numeric(applebalancesheet17[40,3])

#Create table for 
Q_Ratios <- as.data.frame(rbind(Year_2010,Year_2011,Year_2012,Year_2013, Year_2014,Year_2015,Year_2016, Year_2017))

Q_Years <- rbind("2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017")

QR_By_Year <- as.data.frame(cbind(Q_Years, Q_Ratios))

#Created a simple line graph to show the Current Ratio Over time
ggplot(data = QR_By_Year, aes(x = Q_Years, y = V1, group=1)) + 
  geom_line(color="red") +
  geom_hline(aes(yintercept = mean(V1)), color="blue") +
  labs(title="Current Ratios for Apple", x="Year", y="Quick Ratio")

```

##Asset Management Ratios

```{r Inventory Turnover Ratios}
#install the edgar R package
  appleincomestatement <- GetIncome("AAPL", 2010)

print(AnnualReports("AAPL"))  
  
head(GetIncome("GOOG", 2016))

#Liquidity Ratios, Current and Quick.
#Quick Ratio = Current Assets - Inventories/Current Liabilities
  
#Quick Ratio
appleincomestatement10 <- GetIncome("GOOG", 2010)

Year_2010 <- (as.numeric(appleincomestatement10[31,3])/as.numeric(applebalancesheet10[23,3])

Year_2011 <- (as.numeric(applebalancesheet11[31,3])-as.numeric(applebalancesheet11[23,3]))/as.numeric(applebalancesheet11[49,3])

Year_2012 <- (as.numeric(applebalancesheet12[30,3])-as.numeric(applebalancesheet12[22,3]))/as.numeric(applebalancesheet12[48,3])

Year_2013 <- (as.numeric(applebalancesheet13[30,3])-as.numeric(applebalancesheet13[22,3]))/as.numeric(applebalancesheet13[50,3])

Year_2014 <- (as.numeric(applebalancesheet14[30,3])-as.numeric(applebalancesheet14[22,3]))/as.numeric(applebalancesheet14[52,3])

Year_2015 <- (as.numeric(applebalancesheet15[30,3])-as.numeric(applebalancesheet15[22,3]))/as.numeric(applebalancesheet15[54,3])

Year_2016 <- (as.numeric(applebalancesheet16[16,3])-as.numeric(applebalancesheet16[10,3]))/as.numeric(applebalancesheet16[42,3])

Year_2017 <- (as.numeric(applebalancesheet17[16,3])-as.numeric(applebalancesheet17[10,3]))/as.numeric(applebalancesheet17[40,3])

#Create table for 
Q_Ratios <- as.data.frame(rbind(Year_2010,Year_2011,Year_2012,Year_2013, Year_2014,Year_2015,Year_2016, Year_2017))

Q_Years <- rbind("2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017")

QR_By_Year <- as.data.frame(cbind(Q_Years, Q_Ratios))

#Created a simple line graph to show the Current Ratio Over time
ggplot(data = QR_By_Year, aes(x = Q_Years, y = V1, group=1)) + 
  geom_line(color="red") +
  geom_hline(aes(yintercept = mean(V1)), color="blue") +
  labs(title="Current Ratios for Apple", x="Year", y="Quick Ratio")

```