---
title: "Time Value of Money"
author: "Colin James"
date: '2020-07-26'
output:
  html_document:
    df_print: paged
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
lastmod: '2020-07-26T19:09:20-04:00'
categories: []
projects: []
slug: time-value-of-money
subtitle: ''
summary: ''
tags: []
authors: []
---

### Time Value of Money
   Understanding the time value of money is essential to finance.  The value of something is not a static but dynamic thing.  Dating back as far as 5000 BC, this concept is the foundation of modern day finance (https://www.encyclopedia.com/finance/encyclopedias-almanacs-transcripts-and-maps/time-value-money).  Thankfully R makes calculations for TVM easy and efficient.  TVM calculations are broken down into 4 topicsL Future Value, Present Value, Rates of Return and Amortization.
   
```{r}  
#install.packages("modelr")
#install.packages("FinCal")
#install.packages("dplyr")
#install.packages("FinancialMath")

library("modelr")
library("FinCal")
library("dplyr")
library("FinancialMath")
library("ggplot2")

```



## Future Value and Present Value
```{r}
#Below is random cash flow data I have added for these exercises.  We are going to assume our invesment period is 8 years, interest is 10%, and this is an ordinary annunity
cash_flow <- c(-10023,-84949,-84940,-83838,-93838,-73839,-83383,-102939)

i = 1.10
#Manual calculation of FV
(cash_flow[1] * i ^ 7) + (cash_flow[2] * i ^ 6)  +  (cash_flow[3] * i ^ 5) + (cash_flow[4] * i ^ 4) + (cash_flow[5] * i ^ 3) + (cash_flow[6] * i ^ 2) + (cash_flow[7] * i ^ 1) + (cash_flow[8])

#R makes this calculation much easier
fv.uneven(.1,cash_flow)

#FV of Annuity Due.  Int = 10%, T = 8, type = 1 which means payment at the beginning of each period
 fv(.10,8,pv = 0, pmt = 1000, 1)

 #Manual Calculation of PV
 (cash_flow[1]/ (i) ^ 1) + (cash_flow[2] / i ^ 2)  +  (cash_flow[3] / i ^ 3) + (cash_flow[4] / i ^ 4) + (cash_flow[5] / i ^ 5) + (cash_flow[6] / i ^ 6) + (cash_flow[7] / i ^ 7) + (cash_flow[8]
 / i ^ 8)
 
 #Again R makes PV calculation easier
 pv.uneven(.1, cash_flow)
 
 #PV of Annuity Due.  Int = 10%, T = 8, type = 1 which means payment at the beginning of each period
 pv(.1,8, fv = 0, pmt = 1000, type = 1)

```



## Rates of Return
```{r}
#Calculating Rates of Return in R
FV = 149374838
PV = 84933
N = 89
M = 4 # number of compounding periods
Iper = .1

#Calculating annualized returns manually.  Couldnt find a function for this in R
return = (FV/PV) ^ 1/N - 1

return

#Finding the EAR.  
EAR = (1 + Iper/M)^M - 1

EAR

```


#Amortization Tablesin R
```{r}
#Add Numbers to Values to Quick Calculations
FV = 0
PV = 300000
N = 30
M = 12 # frequency of payments per year
Iper = .023


#Step 1 find the PMT
PMT = -pmt(Iper/M,N*M,PV,FV,type = 0)

PMT

#Step 2 create the data frame
Amort <- data.frame(payment = 1:360)

#Step 3 assign beginning value to start loop
Amort[1,2] <- PV

#Step 4 Create for loop to fill in amortization table
 for(i in 1:360){  #number of periods is 360

Amort[i,3] <- Amort[i,2] * Iper/12 #Interest Paid is Beginning Balance x Interest Rate

Amort[i,4] <- PMT - Amort[i,3] #Priciple Paid is equal to the Payment minus Interest Paid

Amort[i,5] <- Amort[i,2] - Amort[i,4] #Ending Balance is equal to the Beginning Balance minus the Principle Paid

Amort[i + 1,2]<- Amort[i,5] #Assign Beginning Balance as last payments Ending Balance

}

#Add column names
names(Amort)[2]<- 'Beginning Balance'
names(Amort)[3]<- 'Interest Paid'
names(Amort)[4]<- 'Principle Paid'
names(Amort)[5]<- 'Ending Balance'

#View the finished table
head(Amort)

#I created this amortization using a for loop, but fortunately R has a package for this

#Amort.table function in R
amort_table <- amort.table(Loan=PV,n=360,i=Iper,ic=12,pf=12)

head(amort_table["Schedule"], n = 10L)

amort_schedule <- as.data.frame(amort_table["Schedule"])
amort_schedule$period <- c(1:360)

#From graph, 
ggplot(data = amort_schedule, aes(x = period, y = Schedule.Principal.Paid)) +
    geom_bar(stat = "identity", color="blue") + 
    labs (title = "Amount of Priciple Paid Over Life of Loan", x = "Payment Period", y = "Dollars")
    
```